---
title: "Cardiovascular Mortality Inequalities in Europe"
author: "Md. Azmir Alam"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
---

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, fig.width = 9, fig.height = 7)
```

## Research question

How do age-standardised cardiovascular mortality rates differ across European countries, sexes and years?

## Load and validate the data

```{r load-data}
data_path <- "data/raw/cardiovascular_mortality.csv"
cvd <- read.csv(data_path, na.strings = c("", "NA"), check.names = FALSE)

required_columns <- c(
  "Country", "Country_code", "Year", "Sex",
  "Age_standardised_rate", "Data_flag", "Source"
)

missing_columns <- setdiff(required_columns, names(cvd))
if (length(missing_columns) > 0) {
  stop("Missing columns: ", paste(missing_columns, collapse = ", "))
}

cvd <- cvd |>
  mutate(
    Year = as.integer(Year),
    Age_standardised_rate = as.numeric(Age_standardised_rate)
  ) |>
  filter(!is.na(Country), !is.na(Year), !is.na(Age_standardised_rate)) |>
  arrange(Country, Year, Sex)

kable(head(cvd), caption = "Eurostat cardiovascular mortality data")
```

## Data-quality summary

```{r quality-summary}
data.frame(
  Measure = c("Rows", "Countries", "First year", "Latest year", "Sex categories"),
  Value = c(
    nrow(cvd),
    n_distinct(cvd$Country),
    min(cvd$Year),
    max(cvd$Year),
    paste(sort(unique(cvd$Sex)), collapse = ", ")
  )
) |>
  kable()
```

## Male–female mortality gap

```{r sex-gap}
sex_gap <- cvd |>
  filter(Sex %in% c("Males", "Females")) |>
  select(Country, Year, Sex, Age_standardised_rate) |>
  pivot_wider(names_from = Sex, values_from = Age_standardised_rate) |>
  mutate(Male_female_gap = Males - Females)

cvd_tableau <- cvd |>
  left_join(sex_gap |> select(Country, Year, Male_female_gap), by = c("Country", "Year"))

latest_gaps <- sex_gap |>
  group_by(Country) |>
  slice_max(Year, n = 1, with_ties = FALSE) |>
  ungroup() |>
  arrange(desc(Male_female_gap)) |>
  slice_head(n = 15)

kable(latest_gaps, digits = 1, caption = "Largest latest male–female mortality gaps")
```

## Countries with the highest latest rates

```{r latest-rate-plot}
latest_total <- cvd |>
  filter(Sex == "Total") |>
  group_by(Country) |>
  slice_max(Year, n = 1, with_ties = FALSE) |>
  ungroup()

highest_rates <- latest_total |>
  slice_max(Age_standardised_rate, n = 15) |>
  arrange(Age_standardised_rate)

mortality_plot <- ggplot(highest_rates, aes(x = Age_standardised_rate, y = reorder(Country, Age_standardised_rate))) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "Highest latest cardiovascular mortality rates",
    x = "Age-standardised rate per 100,000",
    y = "Country"
  ) +
  theme_minimal(base_size = 12)

mortality_plot

dir.create("visuals", showWarnings = FALSE, recursive = TRUE)
ggsave("visuals/cardiovascular_mortality_by_sex.png", mortality_plot, width = 9, height = 7, dpi = 200)
```

## Mortality trends

```{r trends}
selected_countries <- highest_rates |>
  slice_max(Age_standardised_rate, n = 6) |>
  pull(Country)

cvd |>
  filter(Sex == "Total", Country %in% selected_countries) |>
  ggplot(aes(x = Year, y = Age_standardised_rate, colour = Country)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(
    title = "Cardiovascular mortality trends in selected high-rate countries",
    x = "Year",
    y = "Age-standardised rate per 100,000"
  ) +
  theme_minimal(base_size = 12)
```

## Tableau export

```{r tableau-export}
dir.create("data/processed", showWarnings = FALSE, recursive = TRUE)
write.csv(cvd_tableau, "data/processed/tableau_cardiovascular_mortality.csv", row.names = FALSE, na = "")
```

## Interpretation and limitations

Age-standardisation improves comparisons across differently aged populations. However, country reporting practices, coding and mortality-registration completeness still vary. The analysis describes population patterns and should not be interpreted as an individual risk model.
